---
title: "Making nice tables with sparklines"
author: "Jean Clipperton"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(scales)
library(colorspace)
library(gt)
library(tidyverse)
```

## Starting with our data

```{r}
# Let's get our data in order
BA_degrees <- read_csv("data/BA_degrees.csv")

# sort into fields relevant to our inquiry 
BA_degrees_other <- BA_degrees %>%
  mutate(
    field = if_else(field == "Health professions and related programs", "Health professions", field),
    field = fct_other(field,
      keep = c(
        "Business",
        "Health professions",
        "Social sciences and history"
      )
    ),
  ) %>%
  group_by(year, field) %>%
  summarise(pct = sum(pct), .groups = "drop")
BA_degrees_other
```

## Starting with a table
Let's make this a little nicer -- we're going to do some formatting, variable naming, etc. 

```{r echo = FALSE}
BA_degrees_other %>%
  pivot_wider(names_from = field, values_from = pct) %>%
  rename(Year = year) %>% #renaming and going to wide data
  gt() %>%  # getting our table object
  cols_align( # formatting re: alignment and specifying the details
    align = c("right"),
    columns = everything() ) %>%
  cols_width(  Year ~ px(50),
    everything() ~ px(140) ) %>%
  tab_style(
    style = "padding-top:5px;padding-bottom:5px;", #css formatting
    locations = cells_body(columns = everything())  ) %>%
  fmt_percent(  columns = -Year,   decimals = 1 ) %>%
  tab_spanner( label = "Popular Bachelor's degrees over the years",
    columns = everything()  ) %>%
  tab_style(  style = cell_text(weight = "bold"),
    locations = cells_column_spanners() )
```

## Maybe a plot?

```{r degrees-years-plot, echo = FALSE, out.width = "100%", fig.asp = 0.5}
degrees_years_plot <- ggplot(BA_degrees_other, aes(
  x = year, y = pct, group = field, color = field )) +
  geom_line(linewidth = 1, show.legend = FALSE) +
  geom_text(
    data = BA_degrees_other %>% filter(year == min(year)),
    aes(x = year, y = pct,
      label = str_wrap(field, 25)),
    show.legend = FALSE,
    hjust = "right", nudge_x = -1,
    nudge_y = c(-0.02, 0, 0.02, 0),
    fontface = "bold" ) +
  coord_cartesian(clip = "off", xlim = c(1970, 2021)) +
  scale_y_continuous(position = "right", labels = label_percent()) +
  #scale_color_okabeito() +
  theme(plot.margin = margin(0.1, 0.1, 0.1, 1.5, "in")) +
  labs(x = "Year", y = NULL)

degrees_years_plot +
  labs(title = "Popular Bachelor's degrees over the years")
```


## Prep for both:
```{r custom-ggplot, out.width = "40%"}

BA_degrees_other %>%
  pivot_wider(names_from = year, values_from = pct) %>%
  gt() %>%
  cols_align(
    align = "left",
    columns = field ) %>%
  fmt_percent(
    columns = -field, decimals = 0 ) %>%
  cols_label(field = "Field") %>%
  tab_spanner(
    label = "Popular Bachelor's degrees over the years",
    columns = everything()  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_spanners()  )

plot_spark <- function(df) { 
  ggplot(df, aes(x = year, y = pct)) + 
    geom_line(size = 20) + 
    theme_void() } 

BA_degrees_other_plots <- BA_degrees_other %>%
  nest(field_df = c(year, pct)) %>% 
  mutate(plot = map(field_df, plot_spark))

```

## Basic `gt` table with sparklines


```{r sparklines-simple}
BA_degrees_other %>%
  pivot_wider(names_from = year, values_from = pct) %>%
  mutate(ggplot = NA, .after = field) %>%
  gt() %>%
  text_transform( 
    locations = cells_body(columns = ggplot), 
    fn = function(x) { 
      map( .x = BA_degrees_other_plots$plot, .f = ggplot_image, 
        height = px(15), aspect_ratio = 4 ) } ) 
```

### Polishing it up!
Let's turn this into something a little bit nicer. We can add color, turn decimals into percentages, and generally make it a little more professional.



```{r sparklines-polished}
plot_spark_color <- function(df) {
  ggplot(df, aes(x = year, y = pct, color = line_color)) + 
    geom_line(size = 20) +
    theme_void() +
    scale_color_identity() 
}

BA_degrees_other_plots_color <- BA_degrees_other %>%
  mutate(line_color = case_when( 
    field == "Business" ~ "#9D6C06", 
    field == "Health professions" ~ "#077DAA", 
    field == "Social sciences and history" ~ "#026D4E", 
    field == "Other" ~ "#A39A09" 
  )) %>% 
  nest(field_df = c(year, pct, line_color)) %>%
  mutate(plot = map(field_df, plot_spark_color))

BA_degrees_other %>%
  pivot_wider(names_from = year, values_from = pct) %>%
  mutate(ggplot = NA, .after = field) %>%
  gt() %>%
  text_transform(
    locations = cells_body(columns = ggplot),
    fn = function(x) {
      map(BA_degrees_other_plots_color$plot, ggplot_image, height = px(15), aspect_ratio = 4)
    }
  ) %>%
  cols_width(ggplot ~ px(300)) %>%
  cols_align(align = "left", columns = field) %>%
  fmt_percent(columns = where(is.numeric), decimals = 1) %>%
  tab_style(style = cell_text(color = "#9D6C06"), locations = cells_body(rows = 1, columns = field)) %>% 
  tab_style(style = cell_text(color = "#077DAA"), locations = cells_body(rows = 2, columns = field)) %>% 
  tab_style(style = cell_text(color = "#026D4E"), locations = cells_body(rows = 3, columns = field)) %>% 
  tab_style(style = cell_text(color = "#A39A09"), locations = cells_body(rows = 4, columns = field)) %>% 
  cols_label(field = "Field", ggplot = "Trend") %>%
  tab_spanner(label = "Popular Bachelor's degrees over the years", columns = everything()) %>%
  tab_style(style = cell_text(weight = "bold"), locations = cells_column_spanners())
```

## Altering table style
Note that we can change what we want our plot to look like (also note that this particular plot choice revelals the timeframe of our data collectio/reporting). 

```{r sparklines-polished-hist}
plot_spark_color <- function(df) {
  ggplot(df, aes(x = year, y = pct, color = line_color)) + 
    geom_col(size = 10) +
    theme_void() +
    scale_color_identity() 
}

BA_degrees_other_plots_color <- BA_degrees_other %>%
  mutate(line_color = case_when( 
    field == "Business" ~ "#9D6C06", 
    field == "Health professions" ~ "#077DAA", 
    field == "Social sciences and history" ~ "#026D4E", 
    field == "Other" ~ "#A39A09" 
  )) %>% 
  nest(field_df = c(year, pct, line_color)) %>%
  mutate(plot = map(field_df, plot_spark_color))

BA_degrees_other %>%
  pivot_wider(names_from = year, values_from = pct) %>%
  mutate(ggplot = NA, .after = field) %>%
  gt() %>%
  text_transform(
    locations = cells_body(columns = ggplot),
    fn = function(x) {
      map(BA_degrees_other_plots_color$plot, ggplot_image, height = px(15), aspect_ratio = 4)
    }
  ) %>%
  cols_width(ggplot ~ px(300)) %>%
  cols_align(align = "left", columns = field) %>%
  fmt_percent(columns = where(is.numeric), decimals = 1) %>%
  tab_style(style = cell_text(color = "#9D6C06"), locations = cells_body(rows = 1, columns = field)) %>% 
  tab_style(style = cell_text(color = "#077DAA"), locations = cells_body(rows = 2, columns = field)) %>% 
  tab_style(style = cell_text(color = "#026D4E"), locations = cells_body(rows = 3, columns = field)) %>% 
  tab_style(style = cell_text(color = "#A39A09"), locations = cells_body(rows = 4, columns = field)) %>% 
  cols_label(field = "Field", ggplot = "Trend") %>%
  tab_spanner(label = "Popular Bachelor's degrees over the years", columns = everything()) %>%
  tab_style(style = cell_text(weight = "bold"), locations = cells_column_spanners())
```